04_application

### 2020-01-21 chapters/03_methods.Rmd
push new stan to run

Often, the coefficients $\beta_1, ..., \beta_k$ are given simple normal priors such as $\beta_k \sim$ Normal(0, 1) in the case of standardized parameters. When parameter vector $x$ contains many possible covariates, we may reasonably assume that some entries in $\beta$ are 0 -- in other words, that some of the parameters in $x$ have no effect on $Y$. In such a case, it is beneficial to use *sparsity inducing priors* on $\beta$ to ensure that irrelevant parameter coefficients are shrunk to 0.

In frequentist applications, methods such as the LASSO (Least Absolute Shrinkage and Selection Operator) are employed for this type of variable selection. In the Bayesian setting, shrinkage and regularization can be performed using a variety of different priors on $\beta$. **Cite van Erp etc** provide an introduction to many such continuous priors. Ridge regression and the Bayesian LASSO are two of the most approchable sparse regression priors. The horseshoe prior, proposed by **Carvalho et al., 2010**, is another popular choice in Bayesian literature **()** (somehow this accidentally came out sounding exactly like van Erp). Following the notation of **Vehtari**, the horsehoe prior can be written

\begin{align}
\tag{3}
\beta_k \text{ | } \lambda_k, \tau & \sim \text{Normal}(0, \tau^2\lambda_k^2) \\
\lambda_k & \sim \text{Half-Cauchy}(0, 1)
\end{align}

for $k = 1, 2, ..., K$. The horseshoe prior is so-named because for fixed values $\tau = \lambda = 1$, the prior resemebles a Beta(1/2, 1/2) distribution, or a horseshoe.

The horseshoe is often favored because it is a global-local shrinkage prior. This means that $\tau$ shrinks *all* parameters towards 0 while the local parameter $\lambda_k$ and the heavy Cauchy tails allow larger coefficients to remain unshrunk. While this is precisely the goal of a sparsity inducing prior, the horseshoe prior can fail to regularize lrage coefficients *at all*, which can be a problem when parameters are weakly identified by data. Additionally, there is no consensus in the literature for assigning priors to $tau$. **Vehatari..** introduced the *regularized horseshoe* to address both of these shortcomings. The regularized horseshoe builds upon **(3)** such that

\begin{align}
\beta_k \text{ | } \lambda_j, \tau, c & \sim \text{Normal}(0, \tau^2 \tilde{\lambda_k^2}) \\
\tilde{\lambda_k^2} & = \frac{c^2\lambda_k^2}{c^2 + \tau^2\lambda_k^2} \\
\lambda_k & \sim \text{Half-Cauchy}(0, 1) \\
c & \sim ~ \text{Inverse-Gamma}(a, b)
\end{align}

where $c$ > 0 helps to regularize $\beta_k$ far from zero. When $c$ approaches infinity, the regularized horseshoe becomes the standard horseshoe. **This is a place where my understanding is weak**

>>>>

Often, the coefficients $\beta_1, ..., \beta_k$ are given simple normal priors such as $\beta_k \sim$ Normal(0, 1) in the case of standardized parameters. When parameter vector $x$ contains many possible covariates, we may reasonably assume that some entries in $\beta$ are 0 -- in other words, that some of the parameters in $x$ have no effect on $Y$. In frequentist applications, we might use a method such as the LASSO (Least Absolute Shrinkage and Selection Operator) to identify relevant variables.

## Spike and Slab

In the Bayesian setting, there are two primary methods for variable selection. The first is called the "spike-and-slab" prior **(cite)** and has often been considered the "gold standard" for sparse Bayesian regression, or variable selection **(Vehtari)**. This prior is often expressed as

\begin{align}
\tag{3}
\beta_j | \lambda_j, c, \epsilon & \sim \lambda_j \text{Normal}(0, c^2) + (1-\lambda_j)\text{Normal}(0, \epsilon^2) \\
\lambda_j & \sim \text{Bernoulli}(\pi)
\end{align}

for $j = 1, 2, ..., J$ where $\lambda_j \in {0, 1}$ indicates whether $\beta_j$ is from the "spike" (ie, near zero) or the "slab" (ie, nonzero). Priors must be set for the width of the slab $c$ and the prior inclusion probability $\pi$. Here, $\pi$ reflects our prior understanding of the sparsity of $\beta$.

## Horseshoe

The second method for Bayesian sparse regression is to give $\beta$ some continuous *sparsity inducing prior*. **Van Erp et al** provide an introduction to many such continous priors. Ridge regression and the Bayesian LASSO are two of the most approachable priors, but all sparsity inducing priors utilize the same logic: give $\beta$ a prior with *most* of its mass near 0, to shrink irrelevant coefficients, and the rest of its mass "far" from 0. Identifying "far" depends more or less on the specific data and model in question.

The horseshoe prior, proposed by **Carvalho et all, 2010** is a popular choice in Bayesian literature, in part because it is similar to the "gold standard" spike-and-slab. Following the notation of **Vehtari & Pironeen**, the horseshoe prior can be expressed

\begin{align}
\tag{4}
\beta_k \text{ | } \lambda_k, \tau & \sim \text{Normal}(0, \tau^2\lambda_k^2) \\
\lambda_k & \sim \text{Half-Cauchy}(0, 1)
\end{align}

for $k = 1, 2, ..., K$. The horseshoe prior is so-named because for fixed values $\tau = \lambda_k = 1$, the prior resemebles a Beta(1/2, 1/2) distribution, or a horseshoe.

The horseshoe is often favored because it is a global-local shrinkage prior. This means that $\tau$ shrinks *all* parameters towards 0 while the local parameter $\lambda_k$ and the heavy Cauchy tails allow larger coefficients to remain unshrunk. While this is precisely the goal of a sparsity inducing prior, the horseshoe prior can fail to regularize large coefficients *at all*, which can be a problem when parameters are weakly identified by data. Additionally, there is no consensus in the literature for assigning priors to $tau$. **Vehatari..** introduced the *regularized horseshoe* to address both of these shortcomings. The regularized horseshoe builds upon **(3)** such that

\begin{align}
\tag{5}
\beta_k \text{ | } \lambda_j, \tau, c & \sim \text{Normal}(0, \tau^2 \tilde{\lambda_k^2}) \\
\tilde{\lambda_k^2} & = \frac{c^2\lambda_k^2}{c^2 + \tau^2\lambda_k^2} \\
\lambda_k & \sim \text{Half-Cauchy}(0, 1) \\
c & \sim ~ \text{Inverse-Gamma}(a, b)
\end{align}

where $c$ > 0 helps to regularize $\beta_k$ far from zero. When $c$ approaches infinity, the regularized horseshoe becomes the standard horseshoe. **This is a place where my understanding is weak** Using the spike-and-slab as intuition, c corresponds to the width of the slab. **Double check this, but I think** This is, in essence, how we describe "far" from 0.

### 2020-01-21 chapters/04_application.Rmd
add chapters folder
push new stan to run

---
title: "03-methodology"
author: "Raven McKnight"
date: "1/16/2020"
output: html_document
---

@@

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

>>>>

---
title: "04_application"
author: "Raven McKnight"
date: "1/16/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: paper
---

@@

# Case Study

## Background

Metro Transit is the primary transit provider in the Minneapolis-Saint Paul metropolitan area. The agency is interested in predicting transit demand across the region in order to guide route planning and service provision. Currently, Metro Transit uses a set of 5 Transit Market Areas to estimate spatial demand. The Transit Market Areas are writen into the official Transportation Planning Policy (TPP) and are calculated using a simple linear regression. Following the notation of the TPP, the formula for determing Transit Market Areas is expressed

\begin{align}
\text{Transit Market Index }  = & 0.64\text{ * population density} + 0.23\text{ * intersection density} +  \\
& 0.20\text{ * employment_density} + 0.11\text{ * automobile availability}
\end{align}

where each predictor is logged and scaled by developed land acreage per census block group. Here, automobile availability refers to the number of adults over age 16 less the total number of automobiles available in a block group (scaled by developed land acreage). There is an additional indicator variable, omitted in this notation, for the census block group containing the MSP International Airport. For more documentation of the official Transit Market Areas, refere to the TPPs [Appendix G](https://metrocouncil.org/Transportation/Planning-2/Key-Transportation-Planning-Documents/Transportation-Policy-Plan/The-Adopted-2040-TPP-(1)/Final-2040-Transportation-Policy-Plan/2040-TPP-Appendix-G-Transit-Design-and-Perf-Standa.aspx).

Census block groups are split into 5 market areas based on the Transit Market Index described above, where the highest market area (Market Area 1) is expected to support high-frequency, all-day service and the lowest market area (Market Area 5) is expected to support peak commuter express service, if that.

The Transit Market Index values are geographically smoothed to create more-or-less concentric Transit Market Areas.

## Data Description

## Data Preperation

## Modeling

### Poisson Regression

### Horseshoe Prior

### Overdispersion

### Spatial Structure
